---
output:
  md_document:
    variant: markdown_github
---
```{r include=FALSE}
knitr::opts_chunk$set(message=FALSE)
```


# Purpose

This project is the final practical exam for Data Science 871 in which I am required to showcase my understanding and ability of the functional programming paradigm throughout.

### Getting Started

The following code is to create a structured workflow directory as required.

```{r, eval=FALSE}

CHOSEN_LOCATION <- "C:/Users/tianc/Rproj/"
fmxdat::make_project(FilePath = glue::glue("{CHOSEN_LOCATION}"), 
                     ProjNam = "19025831")

Texevier::create_template(directory = glue::glue("C:/Users/tianc/Rproj/19025831/"), template_name = "Question1", open_project = F)
Texevier::create_template(directory = glue::glue("C:/Users/tianc/Rproj/19025831/"), template_name = "Question2", open_project = F)
Texevier::create_template(directory = glue::glue("C:/Users/tianc/Rproj/19025831/"), template_name = "Question3", open_project = F)

```

# Question 1 Solution

The Covid-19 pandemic has wreaked havoc on many facets of our lives, and continue to affect travel, work and how people socially interact. For this question, I visually investigate how different regions globally experienced the pandemic, specifically with reference to the African continent.


## Code used for Figures and Tables

```{r, , message = FALSE, warning = FALSE}

gc() # garbage collection 

#Loading packages

library(pacman)
p_load(tidyverse, lubridate)

# Here I source in all my functions:

list.files('Question1/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

## Data Import

df_owid <- read_csv(glue::glue("Question1/data/Covid/owid-covid-data.csv"))
Continent_list <- c("Europe","North America","South America", "Asia","Africa")
df_owid_adj <- df_owid |> filter(continent %in% Continent_list)

df_death_by_cause <- read_csv(glue::glue("Question1/data/Covid/Deaths_by_cause.csv")) 
df_description <- read_csv("Question1/data/Covid/covid_data_description.csv")

```



The figure below shows the evolution of total positive COVID cases for each continent. Evidently, Africa has significantly less total reported cases than the other continents, with Europe and Asia accounting for the majority. However, a lower level of reported cases in third-world continents is most likely a result of a lack of testing, and can possibly explain why South America also reports such relatively low cases. 


```{r, message = FALSE, warning = FALSE}
plot1 <- cases_per_region(df_owid_adj,xaxis_size = 5, xaxis_rows = 3)

plot1

```

The next figure is similar to above but displays total deaths recorded. Interestingly, even though South America has low levels of cases, the continent displays similar death rates to the of North America and Asia. On the other hand, Africa displays significantly lower deaths than its counterparts.

```{r, message = FALSE, warning = FALSE}

plot2 <- deaths_per_region(df_owid_adj,xaxis_size = 5, xaxis_rows = 3)

plot2

```

One might prematurely assume that the total vaccinations administered in each continent can explain these skewed observations, however, the figure below paints a different picture. At least visually, it seems unlikely that vaccinations account for the widely different experiences between continents, especially between the developed and developing world.

```{r, message = FALSE, warning = FALSE}
plot3 <- vaccinations_per_region(df_owid_adj,xaxis_size = 5, xaxis_rows = 3)

plot3

```


# Question 2 solutions

In this question I attempt to disprove a friend's understanding, contrary to popular belief, that London has good weather that is mostly sunny and not at all cold or rainy. 


```{r, message = FALSE, warning = FALSE}

#Loading packages

library(pacman)
p_load(tidyverse, lubridate)

# Here I source in all my functions:

list.files('Question1/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('Question2/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

# Data

df_weather <- read_csv("Question2/data/London/london_weather.csv", col_types = cols(date = col_date(format = "%Y%m%d")))




```

I first produce the figure below, showcasing the distribution of daily temperatures in London and its average (the red line) from the 1970's through 2020. Notice that the average temperature in the past 40 years hovers around twelve degrees Celsius, with maximum temperature levels vary rarely exceeding 25 degrees. First conclusion: LONDON IS COLD. 

```{r, message = FALSE, warning = FALSE}
g1 <- average_temp_plot(df_weather)

g1

    
```

Secondly, I illustrate in the plot below the distribution of the average London temperature by each season for the last 40 years. Even though Summer temperatures appear pleasant, this pleasantness is diminished by the towering winter distribution, that is heavily skewed towards low temperatures. Moreover, the lowish average temperatures in the Spring and Autumn doesn't exactly invite you to wear short trousers. Therefore, at least to me, this reaffirms the common belief that it is cold at least three quarters of the year in London. 

```{r, message = FALSE, warning = FALSE}

g2 <- Distribution_ave_temp(df_weather, limits = c(-5,30))

g2



```

My friend also, rather incorrectly, stated that London weather is not rainy. Well, the plot below provides the temperature distribution for each month, scattered by the blue dots representing precipitation. It is significantly evident that IT RAINS EVERY MONTH of the year in London. Thereby, completely debunking my friends belief.


```{r, message = FALSE, warning = FALSE}

g3 <- violin_plot(df_weather)

g3

```




# Question 3 solutions

```{r}
#Loading packages

library(pacman)
p_load(tidyverse, lubridate)

# Here I source in all my functions:

list.files('Question1/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))
list.files('Question3/code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))


```





