---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Weather or not to live in London"
subtitle: "Data Science 871 Practical"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Nico Katzke}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Tian Cater"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "University of Stellenbosch" # First Author's Affiliation
Email1: "19025831\\@sun.ac.za" # First Author's Email address

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This article uses graphical simulations to debunk a friends unpopular belief that the weather in London is mostly sunny and not at all cold or rainy. The findings are overwhealmingly indicative that London has mostly cold, cloudy, and rainy wheather throughout the year. This is my solution to question 2 of the practical exam.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

df_weather <- read_csv("data/London/london_weather.csv", col_types = cols(date = col_date(format = "%Y%m%d")))

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

The London weather data used in this article has been kindly provided by the UK National Weather Service, dating back to the 1970's. My argument is structured by referring to three graphical illustrations.


# Lets Get Critical

```{r,  warning =  FALSE, fig.align = 'center', fig.cap = "Daily Average Temperature Since 1979.\\label{Figure1}", fig.ext = 'png', fig.height = 5, fig.width = 7}


g1 <- average_temp_plot(df_weather)

g1


```

Figure \ref{Figure1} illustrated the distribution of the daily temperatures in London from the 1970's through 2020. Notice that the average temperature, represented by the red line, has hovered around twelve degrees Celsius for the past 40 years, with maximum temperature levels vary rarely exceeding 25 degrees. Moreover, the minimum temperatures are regularly below freezing. Thus providing the first conclusion: LONDON IS COLD. 


```{r,  warning =  FALSE, fig.align = 'center', fig.cap = "Temperature Distribution by Season.\\label{Figure2}", fig.ext = 'png', fig.height = 5, fig.width = 7}


g2 <- Distribution_ave_temp(df_weather, limits = c(-5,30))

g2



```

Secondly, I illustrate in figure \ref{2} above the distribution of the average London temperature for each season in the last 40 years. Even though Summer temperatures appear pleasant, this pleasantness is diminished by the towering winter distribution that is heavily skewed towards low temperatures. Moreover, the lowish average temperatures in the Spring and Autumn doesn't exactly invite you to wear short trousers. Therefore, this reaffirms the common belief that it is cold at least three quarters of the year in London.


```{r,  warning =  FALSE, fig.align = 'center', fig.cap = "Temperature and Precipitation Distribution by Month.\\label{Figure3}", fig.ext = 'png', fig.height = 5, fig.width = 7}


g3 <- violin_plot(df_weather)

g3


```

My friend also, rather incorrectly, stated that London weather is not rainy. Well, figure \ref{Figure3} provides the temperature distribution for each month, scattered by the blue dots representing precipitation. It is significantly evident that IT RAINS EVERY MONTH of the year in London. Thereby, completely debunking my friends belief.


