---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "COVID-19:The constrasting experiences of different regions"
subtitle: "Data Science 871 Practical"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Nico Katzke}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Tian Cater"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "University of Stellenbosch" # First Author's Affiliation
Email1: "19025831\\@sun.ac.za" # First Author's Email address

CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This article aims to graphically discuss the constrasting global experiences of the COVID-19 pandemic, specifically when comparing the developed and the developing world. This is my solution to question 1 of the practical exam.
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

list.files('code/', full.names = T, recursive = T) %>% as.list() %>% walk(~source(.))

df_owid <- read_csv(glue::glue("data/Covid/owid-covid-data.csv"))
Continent_list <- c("Europe","North America","South America", "Asia","Africa")
df_owid_adj <- df_owid |> filter(continent %in% Continent_list)

df_death_by_cause <- read_csv(glue::glue("data/Covid/Deaths_by_cause.csv")) 
df_description <- read_csv("data/Covid/covid_data_description.csv")


```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

The Covid-19 pandemic has wreaked havoc by inducing a literal standstill on global activity and disrupting families in all corners of our planet. In its wake, the COVID-19 virus reaffirmed that it does not discriminate between its victims, however, it is apparent that different regions of the globe has had divergent experiences. This article graphically contrasts the pandemics impact on the developed to the that of the developing world by considering the total confirmed COVID-19 cases, its attributed deaths, and  the total vaccinations administered. Special attention is given to the African continent relative to the rest of the world.^[The data used in this article has been kindly provided by OWID (Our
World in Data).]


# Confirmed Cases 

```{r,  warning =  FALSE, fig.align = 'center', fig.cap = "Total confirmed cases of COVID-19 per Region.\\label{Figure1}", fig.ext = 'png', fig.height = 5, fig.width = 7}

  # Source the plot with a function, and then print it as below.
    # Going through the function, you will note that I'm able to source other functions in the code folder in these functions too.
# That's the beauty of centralizing your functions and loading it in the top.

plot1 <- cases_per_region(df_owid_adj,xaxis_size = 5, xaxis_rows = 3)

plot1

```

Figure \ref{Figure1} above shows the evolution of total reported COVID cases for each continent. Evidently, Africa has significantly less reported cases than the other continents, with Europe and Asia accounting for the majority. South America also showcases substantially lower cases, hinting that the developing world was less vulnerable with regards to the virus.  However, importantly, a lower level of reported cases in third-world continents is very likely a result of a lack of COVID testing.

# Deaths Attributed to COVID-19

```{r,  warning =  FALSE, fig.align = 'center', fig.cap = "Total Deaths Attributed to COVID-19 per Region.\\label{Figure2}", fig.ext = 'png', fig.height = 6, fig.width = 6}

  

 plot2 <- deaths_per_region(df_owid_adj,xaxis_size = 5, xaxis_rows = 3)

plot2
```

When considering the total deaths attributed to COVID per region in figure \ref{Figure2} above, some interesting results arise; Notwithstanding South America's significantly lower levels of cases in figure \ref{Figure1}, the continent displays similar death rates to that of North America and Asia. On the one hand, this result makes sense as infected individuals in developing nations do not have similar accessibility to health care. However, on the other hand, when comparing South American death numbers to that of Africa's, the health-care argument loses its power. Maybe in considering the total number of vaccines administered in figure \ref{Figure3} below a clearer explanation will reveal itself. 


# Vaccinations Administered

```{r, warning =  FALSE, fig.align = 'center', fig.cap = "Total number of COVID-19 vaccination doses administered per Region.\\label{Figure3}", fig.ext = 'png', fig.height = 6, fig.width = 6}

plot3 <- vaccinations_per_region(df_owid_adj,xaxis_size = 5, xaxis_rows = 3)

plot3

```

One might prematurely assume that the total vaccinations administered in each continent can explain the skewed observations in figures \ref{Figure1} and \ref{Figure2}, however, the figure \ref{Figure3} below paints a different picture. At least visually, it seems unlikely that vaccinations account for the widely different experiences between continents, especially between the developed and developing world. However, it will be wise to refrain from drawing conclusions based on the admittedly non-robust data analysis conducted here. Further econometric analysis needs to be conducted.   